---
title: "Understanding DAGs"
author: "Anders Sundelin"
date: "2022-12-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dagitty)
library(ggdag)
library(ggplot2)
#library(rethinking)
```

# Getting down to DAGs

Initial DAG:

```{r }
introd.dag <- dagitty("dag {
                      AUTHOR -> INTROD;
                      REPO -> INTROD;
                      TEAM -> INTROD;
                      REASON -> INTROD;
                      ADDED -> INTROD;
                      ISSUES -> INTROD;
                      COMPLEX -> INTROD;
                      FILE -> ISSUES;
                      FILE -> REPO;
                      FILE -> COMPLEX;
                      AUTHOR -> TEAM;
                      AUTHOR -> REPO;
                      TEAM -> REPO;
                      }")
coordinates(introd.dag) <- list(x=c(COMPLEX=0, FILE=1, ISSUES=1, ADDED=-1, INTROD=1, REPO=2, REASON=-1, AUTHOR=0, TEAM=2),
                                y=c(COMPLEX=0, FILE=0, ISSUES=1, ADDED=1 , INTROD=2, REPO=2, REASON=3 , AUTHOR=3, TEAM=3))

drawdag(introd.dag)
```

Explanation of the variables in the DAG:

* INTROD - this is the thing we are trying to model/predict. It is the number of *introduced duplicates* in a particular file, in a particular change (commit). Zero or larger (absolute upper bound is on the order of the total number of lines in all files - due to the way Sonar calculates duplicates, though this assumes duplicates are counted line per line - in practice the tool will cluster identical lines or modules)
* ADDED  - Number of added lines in the particular file, in the particular commit, as measured via Git (--numstat)
* REASON - The reason for changing the file - factor levels include "Added functionality", "TR fix", "Refactoring/spontaneous improvement".
* AUTHOR - The author that started the change (Git Author).
* TEAM   - The team that the author was part of, at the time of the commit first being created.
* REPO   - The repository holding the files. Authors and teams are more or less familiar with individual repositories (depending on whether they have worked there before).
* ISSUES - The number of existing duplicates, in the file, as measured by Sonar, before the commit is applied.
* FILE   - The changed file. Files always belong to a repo. So, given that I know what file it is, I know what repo is affected. But a repo has (most likely) many files.
* COMPLEX - The McCabe complexity of the file (as calculated by Sonar), before the change was made.

Causal reasoning:

* By modifying code, authors make changes to existing files, create new files, or delete existing files.
* Authors belong to a team, interact with team members (reviews, merges, discusses changes).
* Both authors and teams have various levels of understanding of code repositories.
* Repositories serve as the grouping level of files. Each file belong to one, and only one, repository, and if I know what file is affected, I also know what repository was affected. But the converse is not true.
* Individual files have individual number of existing issues (duplicates), and existing McCabe complexity. Both values start at 0, and are bounded by the number of lines in the file (at least same magnitude).

The purpose of the model is to ascertain how the model can be used to illustrate and predict "team behaviour", in particular in an ownership context.
Teams that "care for" their repo are likely to introduce less duplicates than other, less caring (or more junior) teams.


Questions:

* We could label the arrow between AUTHOR and REPO with "familiarity". How do we reason about the direction of the arrow? Should the causal effect flow "from AUTHOR, via REPO, to INTROD", or is it "from REPO, via AUTHOR, to INTROD"? 
* Similar discussion for TEAM.
* This is of course super important, as it determines whether REPO is a collider or not.
* We could put ADDED and REASON inside a grouping level called COMMIT, as they are effectively tied together there. But it is unclear what that would give us - we are not comparing commits per se.



```{r}
adjustmentSets(introd.dag, exposure = c("TEAM", "ADDED", "COMPLEX", "ISSUES", "REASON"), outcome = "INTROD")
```

So, if I were to include TEAM, ADDED, COMPLEX, ISSUES and REASON in my causal model, I also need to include both AUTHOR and FILE. But not REPO (due to the collider bias).


If I also include "REPO":

```{r}
adjustmentSets(introd.dag, exposure = c("TEAM", "ADDED", "COMPLEX", "ISSUES", "REASON", "REPO"), outcome = "INTROD")
```

So, if I include REPO in my predictors, then I should include AUTHOR, but not FILE.


```{r}
impliedConditionalIndependencies(introd.dag)
```


# Modified DAG

If we instead reason that AUTHORs only introduces duplicates via FILEs, and not directly influences INTROD, then our DAG becomes:

```{r}
introd.dag.2 <- dagitty("dag {
                      AUTHOR -> FILE;
                      REPO -> INTROD;
                      TEAM -> INTROD;
                      REASON -> INTROD;
                      ADDED -> INTROD;
                      ISSUES -> INTROD;
                      COMPLEX -> INTROD;
                      FILE -> ISSUES;
                      FILE -> REPO;
                      FILE -> COMPLEX;
                      AUTHOR -> TEAM;
                      TEAM -> REPO;
                      }")
coordinates(introd.dag.2) <- list(x=c(COMPLEX=0, FILE=1, ISSUES=1, ADDED=-1, INTROD=1, REPO=2, REASON=-1, AUTHOR=2, TEAM=2),
                                y=c(COMPLEX=0, FILE=0, ISSUES=1, ADDED=1 , INTROD=2, REPO=3, REASON=3 , AUTHOR=0, TEAM=2))

drawdag(introd.dag.2)

```



```{r}
adjustmentSets(introd.dag.2, exposure = c("TEAM", "COMPLEX", "ISSUES", "ADDED", "REASON"), outcome = "INTROD")
```

## Criticism of the above model

Important to note that arrows in a DAG should refer to _causal_ relationships. A _causes_ B, not "A is part of B", or "A is an attribute of B".
Suggestion from mentor is to make a matrix (spreadsheet) with motivations, including arrows and direction for all plausible combinations.

But be sure differentiate between causal effects and correlations or hierarchical information.

With that said, we will revise the above DAGs. 
AUTHOR has an effect on ADDED, but FILE does not have a causal relationship with REPO (though given that I know what file is affected, the repo is also given - but this is a correlation, not _causation_).

Suggestion from mentor is to model the relationships in a matrix form, write down all possible combinations, and also indicate the direction of the arrows.

In particular, given a DAG and an analysis, one may also perform sensitivity analysis (akin to physics' pertubation analysis), that is, reasoning about how many or how strong unobserved counfounds that would be needed to erase a particular effect. https://evalf21.classes.andrewheiss.com/example/confounding-sensitivity/


* AUTHOR (or, COMMITTER, as this is the person doing the final merge) would have a causal effect on ADDED.
* FILE would not have a causal effect on REPO, but either (or both) would have a causal effect on AUTHOR or COMMITTER. 
  People are more or less famililar with the functions, tests, structure of particular FILEs or REPOs (this is the ownership part).
* A FILE will _have_ an existing complexity (COMPLEX), but a particular FILE would not _cause_ that same COMPLEX. It is still plausible that COMPLEX would affect the AUTHOR (and/or TEAM) via causation (harder to make changes to complex files).
* In itself, neither FILE nor REPO would affect INTROD. The unfamiliarity would come through either AUTHOR/COMMITTER/TEAM.

This would indicate that if I want to predict INTROD based on the TEAM, COMPLEX, ISSUES, ADDED and REASON variables, I also need to include FILE in my model. But not REPO or AUTHOR.

```{r}
adjustmentSets(introd.dag.2, exposure = c("TEAM", "COMPLEX", "ISSUES", "ADDED", "REASON", "REPO"), outcome = "INTROD")
```

No further parameters should be added if my linear regression includes TEAM, COMPLEX, ISSUES, ADDED, REASON and REPO.


```{r}
adjustmentSets(introd.dag.2, exposure = c("AUTHOR", "COMPLEX", "ISSUES", "ADDED", "REASON"), outcome = "INTROD")
```

There does not seem to be any adjustmentSets for these parameters. Using FILE instead of ISSUES or COMPLEX will solve the problem, returning an empty set again.


```{r}
impliedConditionalIndependencies(introd.dag.2)
```


```{r}
introd.dag.3 <- dagitty("dag {
                      AUTHOR [exposure];
                      TEAM [exposure];
                      INTROD [outcome];
                      AUTHOR -> ADDED;
                      AUTHOR -> REMOVED;
                      TEAM -> AUTHOR;
                      TEAM -> ADDED;
                      TEAM -> REMOVED;
                      FILE -> AUTHOR;
                      FILE -> TEAM;
                      REPO -> AUTHOR;
                      REPO -> TEAM;
                      ISSUES -> AUTHOR;
                      ISSUES -> TEAM;
                      COMPLEX -> AUTHOR;
                      COMPLEX -> TEAM;
                      ADDED -> INTROD;
                      REASON -> AUTHOR;
                      REASON -> TEAM;
                      REMOVED -> INTROD;
                      }")
#coordinates(introd.dag.2) <- list(x=c(COMPLEX=0, FILE=1, ISSUES=1, ADDED=-1, INTROD=1, REPO=2, REASON=-1, AUTHOR=2, TEAM=2),
#                                y=c(COMPLEX=0, FILE=0, ISSUES=1, ADDED=1 , INTROD=2, REPO=3, REASON=3 , AUTHOR=0, TEAM=2))

drawdag(introd.dag.3)

```

```{r}
impliedConditionalIndependencies(introd.dag.3)
```

```{r}
adjustmentSets(introd.dag.3, exposure = c("AUTHOR", "TEAM", "COMPLEX", "ADDED", "ISSUES", "REPO"), outcome = "INTROD")
```

```{r}
adjustmentSets(introd.dag.3, exposure = c("AUTHOR"), outcome = "INTROD")
```

So, this tells us that if we want to examine the total way that TEAM and AUTHOR influences INTROD, we should also include REASON into our model.
This is becuase of the backdoor path through REMOVED. We assume that REASON will have an impact at least through the removal of files.

The model can be more graphically explored on http://www.dagitty.net/
This also includes coloring, and using node types, making the DAG a wee bit more understandable.

Would AUTHOR have *any* effect on INTROD except through ADDED or REMOVED? I think not? Likewise for TEAM... They might influence Author (committer, really).

# Only action through added/removed lines

We could plausibly argue that the only action an AUTHOR or a TEAM has on the number of issues is through the number of ADDED or REMOVED lines. These are, after all, the sole change that the AUTHOR (or other users) have on the file in question.

Does it change any inference?

```{r}
introd.dag.4 <- dagitty("dag {
                      AUTHOR [exposure];
                      TEAM [exposure];
                      INTROD [outcome];
                      AUTHOR -> ADDED;
                      AUTHOR -> REMOVED;
                      TEAM -> AUTHOR;
                      FILE -> AUTHOR;
                      FILE -> TEAM;
                      REPO -> AUTHOR;
                      REPO -> TEAM;
                      ISSUES -> AUTHOR;
                      ISSUES -> TEAM;
                      COMPLEX -> AUTHOR;
                      COMPLEX -> TEAM;
                      ADDED -> INTROD;
                      REASON -> AUTHOR;
                      REASON -> REMOVED;
                      REMOVED -> INTROD;
                      }")
#coordinates(introd.dag.4) <- list(x=c(COMPLEX=0, FILE=1, ISSUES=1, ADDED=-1, INTROD=1, REPO=2, REASON=-1, AUTHOR=2, TEAM=2),
#                                y=c(COMPLEX=0, FILE=0, ISSUES=1, ADDED=1 , INTROD=2, REPO=3, REASON=3 , AUTHOR=0, TEAM=2))

drawdag(introd.dag.4)
```

```{r}
adjustmentSets(introd.dag.4, exposure = c("TEAM", "AUTHOR"), outcome="INTROD")
```

```{r}
adjustmentSets(introd.dag.4, exposure = c("TEAM", "AUTHOR", "COMPLEX", "ISSUES", "REPO"), outcome="INTROD")
```

```{r}
impliedConditionalIndependencies(introd.dag.4)
```

We could add a COMMIT node, to indicate that both AUTHOR and TEAM influence commits, and commits cause ADDED or REMOVED lines.

```{r}
introd.dag.5 <- dagitty("dag {
                      AUTHOR [exposure];
                      TEAM [exposure];
                      INTROD [outcome];
                      COMMIT -> ADDED;
                      COMMIT -> REMOVED;
                      AUTHOR -> COMMIT;
                      TEAM -> COMMIT;
                      FILE -> AUTHOR;
                      FILE -> TEAM;
                      REPO -> AUTHOR;
                      REPO -> TEAM;
                      ISSUES -> AUTHOR;
                      ISSUES -> TEAM;
                      COMPLEX -> AUTHOR;
                      COMPLEX -> TEAM;
                      ADDED -> INTROD;
                      REASON -> AUTHOR;
                      REASON -> REMOVED;
                      REMOVED -> INTROD;
                      }")
#coordinates(introd.dag.4) <- list(x=c(COMPLEX=0, FILE=1, ISSUES=1, ADDED=-1, INTROD=1, REPO=2, REASON=-1, AUTHOR=2, TEAM=2),
#                                y=c(COMPLEX=0, FILE=0, ISSUES=1, ADDED=1 , INTROD=2, REPO=3, REASON=3 , AUTHOR=0, TEAM=2))

drawdag(introd.dag.5)

```

```{r}
adjustmentSets(introd.dag.5, exposure = c("AUTHOR", "TEAM", "COMPLEX", "REPO", "ISSUES"), outcome = "INTROD")
```

Adding COMMIT as a descendent of AUTHOR (and before ADDED and REMOVED) does not change the number of nodes that should be added to our graph. We should still add REASON, to close the backdoor path through REMOVED (i.e. the possibility that refactorings may bias the results, related to removed files).

Adding committer (as the person finally making the change to the repo). Committer and author MAY be the same person, but does not have to. Only the commiter influences the final result.

```{r}
introd.dag.6 <- dagitty("dag {
                      AUTHOR [exposure];
                      TEAM [exposure];
                      COMMITTER [exposure]
                      INTROD [outcome];
                      COMMITTER -> ADDED;
                      COMMITTER -> REMOVED;
                      AUTHOR -> COMMITTER;
                      TEAM -> COMMITTER;
                      FILE -> AUTHOR;
                      FILE -> TEAM;
                      REPO -> AUTHOR;
                      REPO -> TEAM;
                      ISSUES -> AUTHOR;
                      ISSUES -> TEAM;
                      COMPLEX -> AUTHOR;
                      COMPLEX -> TEAM;
                      ADDED -> INTROD;
                      REASON -> AUTHOR;
                      REASON -> REMOVED;
                      REMOVED -> INTROD;
                      }")
#coordinates(introd.dag.4) <- list(x=c(COMPLEX=0, FILE=1, ISSUES=1, ADDED=-1, INTROD=1, REPO=2, REASON=-1, AUTHOR=2, TEAM=2),
#                                y=c(COMPLEX=0, FILE=0, ISSUES=1, ADDED=1 , INTROD=2, REPO=3, REASON=3 , AUTHOR=0, TEAM=2))

drawdag(introd.dag.6)
```

```{r}
adjustmentSets(introd.dag.6, exposure = c("AUTHOR", "TEAM", "COMMITTER", "COMPLEX", "REPO", "ISSUES"), outcome = "INTROD")

```

Having committer in the DAG does not change anything.

```{r}
adjustmentSets(introd.dag.6, exposure = c("TEAM", "COMMITTER", "COMPLEX", "REPO", "ISSUES"), outcome = "INTROD")
```

If we swap AUTHOR for COMMITTER (person doing the final change), we still have to include REASON in our model. But having AUTHOR there is not needed, if we are only striving to assess COMMITTER

# Team-only model

We could model the author&team as one entity, due to our sparse data. That effectively collapses the interaction and questions what is the causal relation between authors and teams (e.g. does the author cause the team's behaviour, or vice versa).

```{r}
introd.dag.7 <- dagitty("dag {
                      TEAM [exposure];
                      REPO [exposure]
                      INTROD [outcome];
                      TEAM -> ADDED;
                      TEAM -> REMOVED;
                      FILE -> TEAM;
                      REPO -> TEAM;
                      ISSUES -> TEAM;
                      COMPLEX -> TEAM;
                      ADDED -> INTROD;
                      REASON -> TEAM;
                      REMOVED -> INTROD;
                      }")
drawdag(introd.dag.7)
```

Adjustment set is empty, as all causal paths flow through TEAM. We could of couse still investigate the impact of ISSUES, REASON, COMPLEX.

If we were to believe that some reasons (e.g. improvements or refactorings) would impact added lines, then that would become a collider and should be part of the model, even if it is not an outcome per se.

```{r}
introd.dag.8 <- dagitty("dag {
                      TEAM [exposure];
                      REPO [exposure]
                      INTROD [outcome];
                      TEAM -> ADDED;
                      TEAM -> REMOVED;
                      FILE -> TEAM;
                      REPO -> TEAM;
                      ISSUES -> TEAM;
                      COMPLEX -> TEAM;
                      ADDED -> INTROD;
                      REASON -> TEAM;
                      REASON -> ADDED;
                      REMOVED -> INTROD;
                      }")
drawdag(introd.dag.8)
```

But it seems unlikely that REASON would impact ADDED lines without first going through ADDED or REMOVED lines (e.g. does the REASON has a different cause than going through the AUTHOR/TEAM process?)

```{r}
adjustmentSets(introd.dag.8, exposure = "REPO", outcome = "INTROD", effect = "direct")
```

### New causal model

```{r}
testImplications <- function( covariance.matrix, sample.size ){
	library(ggm)
	tst <- function(i){ pcor.test( pcor(i,covariance.matrix), length(i)-2, sample.size )$pvalue }
tos <- function(i){ paste(i,collapse=" ") }
implications <- list(c("COMPLEX","FEEDBACK"),
		c("COMPLEX","REQ"),
		c("COMPLEX","NCOMMITS"),
		c("DUP","FEEDBACK"),
		c("DUP","REQ"),
		c("DUP","NCOMMITS"),
		c("FEEDBACK","INTROD","ADD","COMPLEX","DEL","DUP"),
		c("FEEDBACK","REQ"),
		c("FEEDBACK","TEAM_REPO"),
		c("FEEDBACK","NCOMMITS"),
		c("INTROD","REQ","DUP","COMPLEX","DEL","ADD"),
		c("INTROD","TEAM_REPO","DEL","ADD","DUP","COMPLEX"),
		c("INTROD","NCOMMITS","DUP","COMPLEX","DEL","ADD"),
		c("REQ","TEAM_REPO"),
		c("REQ","NCOMMITS"),
		c("TEAM_REPO","NCOMMITS"))
	data.frame( implication=unlist(lapply(implications,tos)),
		pvalue=unlist( lapply( implications, tst ) ) )

}

introd.dag.9 <- dagitty("dag {
bb=\"-4.944,-4.291,4.847,5.173\"
ADD [pos=\"0.918,-0.890\"]
CARE [latent,pos=\"-2.168,0.674\"]
COMPLEX [pos=\"-2.590,-1.081\"]
DEL [pos=\"-1.062,-0.699\"]
DUP [pos=\"-1.826,-0.777\"]
EXP [latent,pos=\"-1.215,0.787\"]
FEEDBACK [pos=\"1.355,0.179\"]
INTROD [outcome,pos=\"-0.995,-2.778\"]
KNOW [latent,pos=\"0.428,0.919\"]
NCOMMITS [pos=\"1.165,1.326\"]
REQ [pos=\"0.923,-1.885\"]
TEAMREPO [exposure,pos=\"-1.262,1.688\"]
ADD -> INTROD
CARE -> COMPLEX
CARE -> DUP
CARE -> EXP
COMPLEX -> INTROD
DEL -> INTROD
DUP -> INTROD
EXP -> KNOW
FEEDBACK -> ADD
FEEDBACK -> DEL
FEEDBACK -> KNOW
KNOW -> ADD
KNOW -> DEL
NCOMMITS -> KNOW
REQ -> ADD
REQ -> DEL
TEAMREPO -> CARE
TEAMREPO -> KNOW
                      }")
drawdag(introd.dag.9)
```
```{r}
ggdag(introd.dag.9) + theme_dag_blank()
```

The backdoor criterion says we should exclude controls that are descendents of the treatment along paths to the outcome.
If our target is the Average Causal Effect, we want to leave all channels through which the causal effect flows untouched.
We should not control for mediators.
Neutral controls might work though. That is, when Z is a cause of the mediator, but not an effect of the treatment.

```{r}
introd.dag.10 <- dagitty("dag {
bb=\"-6.094,-5.092,6.372,5.521\"
ADD [pos=\"-0.796,0.553\"]
CARE [latent,pos=\"-1.863,-0.920\"]
COMPLEX [pos=\"-2.299,-2.207\"]
DEL [pos=\"-2.339,0.520\"]
DUP [pos=\"-2.902,-2.212\"]
EXP [latent,pos=\"-1.159,-0.920\"]
FEEDBACK [pos=\"1.121,-0.942\"]
INTROD [outcome,pos=\"-1.400,-3.971\"]
KNOW [latent,pos=\"-0.106,-0.931\"]
NCOMMITS [pos=\"-1.112,-1.558\"]
NTESTS [pos=\"1.155,-1.574\"]
REQ [exposure,pos=\"-1.782,1.553\"]
ADD -> CARE
ADD -> KNOW
CARE -> COMPLEX
CARE -> DUP
CARE -> INTROD
COMPLEX -> INTROD
DEL -> CARE
DEL -> KNOW
DUP -> INTROD
EXP -> KNOW
FEEDBACK -> ADD
FEEDBACK -> DEL
FEEDBACK -> INTROD
FEEDBACK -> KNOW
KNOW -> INTROD
NCOMMITS -> EXP
NTESTS -> FEEDBACK
REQ -> ADD
REQ -> DEL
}")
drawdag(introd.dag.10)
```

```{r}
adjustmentSets(introd.dag.10)
```

```{r}
introd.dag.11 <- dagitty("dag {
bb=\"-6.094,-5.092,6.372,5.521\"
ADD [exposure,pos=\"-0.636,-1.651\"]
CARE [latent,pos=\"-2.527,0.459\"]
COMPLEX [pos=\"-2.299,-2.207\"]
DEL [exposure,pos=\"-1.507,-1.662\"]
DUP [pos=\"-2.902,-2.212\"]
EXP [latent,pos=\"-0.159,0.184\"]
FEEDBACK [pos=\"-1.004,-0.503\"]
INTROD [outcome,pos=\"-1.400,-3.971\"]
KNOW [latent,pos=\"0.484,-1.410\"]
NCOMMITS [pos=\"-0.837,0.717\"]
NTESTS [pos=\"-1.601,0.151\"]
ADD -> INTROD
CARE -> ADD
CARE -> COMPLEX
CARE -> DEL
CARE -> DUP
COMPLEX -> INTROD
DEL -> INTROD
DUP -> INTROD
EXP -> KNOW
FEEDBACK -> ADD
FEEDBACK -> DEL
FEEDBACK -> KNOW
KNOW -> INTROD
NCOMMITS -> EXP
NTESTS -> FEEDBACK
}
")
drawdag(introd.dag.11)

```

```{r}
introd.dag.12 <- dagitty("
             dag {
bb=\"-3.165,-3.602,3.813,3.596\"
ADD [exposure,pos=\"-1.894,-2.371\"]
COMPLEX [pos=\"0.784,2.997\"]
DUP [pos=\"3.232,0.427\"]
FEAR [latent,pos=\"0.028,0.052\"]
INTROD [outcome,pos=\"0.529,-3.002\"]
TESTS [pos=\"-2.584,1.942\"]
ADD -> INTROD
COMPLEX -> FEAR
DUP -> FEAR
FEAR -> ADD
FEAR -> INTROD
TESTS -> FEAR
}
                         ")
drawdag(introd.dag.12)

```

```{r}
adjustmentSets(introd.dag.12)
```


```{r}
introd.dag.13 <- dagitty("dag {
bb=\"-3.165,-3.602,3.813,3.596\"
ADD [exposure,pos=\"-1.894,-2.371\"]
COMPLEX [pos=\"0.313,-1.217\"]
DUP [pos=\"1.504,-1.362\"]
INTROD [outcome,pos=\"0.529,-3.002\"]
TESTS [pos=\"-0.975,-1.011\"]
ADD -> INTROD
COMPLEX -> ADD
COMPLEX -> INTROD
DUP -> ADD
DUP -> INTROD
TESTS -> ADD
TESTS -> INTROD
}

                         ")
drawdag(introd.dag.13)

```

```{r}
adjustmentSets(introd.dag.13)
```


We have four possible adjustment sets. We have to condition on X4, but as it is a collider (receiving arrows both from X1 and X2), this will open the causal path, introducing bias.

```{r}
pearl.dag.fig.3.4 <- dagitty("dag {
Xi [exposure]
Xj [outcome]
X1 -> X3
X1 -> X4
X2 -> X4 
X2 -> X5
X3 -> Xi
X4 -> Xi
X4 -> Xj
X5 -> Xj
Xi -> X6
X6 -> Xj
                             }")
drawdag(pearl.dag.fig.3.4)
adjustmentSets(pearl.dag.fig.3.4)
```
If we make X4 a pipe instead of a collider, it is enough to condition on X4.

```{r}
pearl.dag.fig.3.4.pipe <- dagitty("dag {
Xi [exposure]
Xj [outcome]
X1 -> X3
X1 -> X4
X4 -> X2 
X2 -> X5
X3 -> Xi
X4 -> Xi
X4 -> Xj
X5 -> Xj
Xi -> X6
X6 -> Xj
                             }")
drawdag(pearl.dag.fig.3.4.pipe)
adjustmentSets(pearl.dag.fig.3.4.pipe)
```

```{r}
my.dag <- dagitty("dag {
ADD [exposure]
INTROD [outcome]
ADD -> REM
ADD -> DUP
ADD -> COMPLEX 
DUP -> COMPLEX
REM -> INTROD
DUP -> INTROD
COMPLEX -> INTROD
ADD -> INTROD
}")
drawdag(my.dag)
adjustmentSets(my.dag)

```

```{r}
d.0 <- dagitty("dag {
lang [exposure]
rank [outcome]
skill [unobserved]
lang -> rank
lang -> size
nick -> lang 
nick -> size
nick -> skill
challenge -> nick
challenge -> rank
challenge -> skill
skill -> rank
}")
drawdag(d.0)
adjustmentSets(d.0)
impliedConditionalIndependencies(d.0)
```

```{r}
d.1 <- dagitty("dag {
lang [exposure]
rank [outcome]
skill [unobserved]
lang -> rank
lang -> size
nick -> lang 
nick -> size
nick -> skill
challenge -> nick
challenge -> rank
challenge -> skill
skill -> rank
skill -> size
}")
drawdag(d.1)
adjustmentSets(d.1)
impliedConditionalIndependencies(d.1)
```

```{r}
d.2 <- dagitty("dag {
lang [exposure]
rank [outcome]
skill [unobserved]
lang -> rank
lang -> size
nick -> lang 
nick -> size
nick -> skill
challenge -> nick
challenge -> rank
challenge -> skill
skill -> rank
skill -> size
nick -> size
}")
drawdag(d.2)
adjustmentSets(d.2)
impliedConditionalIndependencies(d.2)
```

```{r}
test.dag <- dagitty("dag {
ADD [exposure]
INTROD [outcome]
KNOW [unobserved]
CLEAN [unobserved]
CLEAN -> REM
CLEAN -> ADD
REM -> INTROD
DUP -> COMPLEX
KNOW -> COMPLEX
KNOW <- DUP
KNOW <- ADD
DUP -> ADD
COMPLEX -> ADD
ADD -> INTROD
}")
drawdag(test.dag)
adjustmentSets(test.dag)
impliedConditionalIndependencies(test.dag)

```
```{r}
test.dag <- dagitty("dag {
ADD [exposure]
INTROD [outcome]
KNOW [unobserved]
CLEAN [unobserved]
CLEAN -> REM
CLEAN -> ADD
REM -> INTROD
COMPLEX -> INTROD
DUP -> INTROD
DUP -> COMPLEX -> KNOW
DUP -> KNOW
KNOW -> ADD
DUP -> ADD
COMPLEX -> ADD
ADD -> INTROD
}")
ggdag(test.dag)
#drawdag(test.dag)
adjustmentSets(test.dag)
impliedConditionalIndependencies(test.dag)

```


```{r}
adjustmentSets(test.dag)

```


```{r}
updated_dag <- dagitty("dag {
bb=\"0,0,1,1\"
ADD [adjusted,pos=\"0.603,0.269\"]
CLEAN [latent,pos=\"0.186,0.482\"]
COMP [adjusted,pos=\"0.529,0.397\"]
DEL [adjusted,pos=\"0.594,0.490\"]
DUP [adjusted,pos=\"0.512,0.103\"]
INTR [outcome,pos=\"0.928,0.296\"]
KNOW [latent,pos=\"0.198,0.067\"]
TEAM [exposure,pos=\"0.065,0.279\"]
ADD -> INTR
CLEAN -> ADD
CLEAN -> DEL
COMP -> ADD
COMP -> INTR
COMP -> KNOW
DEL -> ADD
DEL -> INTR
DUP -> ADD
DUP -> COMP
DUP -> INTR
DUP -> KNOW
KNOW -> ADD
TEAM -> ADD
TEAM -> CLEAN
TEAM -> COMP
TEAM -> DEL
TEAM -> DUP
TEAM -> KNOW
}") |> tidy_dagitty()
ggdag(updated_dag, exposures = "TEAM") + theme_dag()
```
```{r}
coords <- data.frame(name=c("TEAM", "KNOW", "CLEAN", "DUP", "COMP", "ADD", "DEL", "INTR"), 
                     x=c(        0,    0.2,    0.4,   0.4,   0.35,   0.55,   0.38, 1.0),
                     y=c(     -0.03,    0.2,   -0.08,   0.2,  0.03,   0.07,  -0.2, 0))
dagified <- dagify( INTR ~ DEL + COMP + ADD + DUP + TEAM, 
                    COMP ~ TEAM + DUP,
                    DEL ~ TEAM + CLEAN,
                    ADD ~ COMP + CLEAN + KNOW + DUP,
                    DUP ~ TEAM,
                    CLEAN ~ TEAM,
                    KNOW ~ TEAM + DUP + COMP,
                    exposure = "TEAM",
                    outcome = "INTR",
                    latent = c("CLEAN", "KNOW"),
                    coords=coords)
(p <- ggdag_status(dagified) + theme_dag() + theme(legend.position="bottom") + guides(color = guide_legend(override.aes = list(size = 2), title=NULL, direction="horizontal", nrow = 1)) 
  )
ggsave("causal-dag.pdf", p, device = "pdf", dpi = 1200, width=15, height=17.8, units="cm") 
```

```{r}
impliedConditionalIndependencies(dagified)
```


```{r}
(p <- ggdag_adjustment_set(dagified, shadow = T) + theme_dag() )

```
```{r}
coords <- data.frame(name=c("TEAM", "KNOW", "CLEAN", "DUP", "COMP", "ADD", "REM", "INTR"), 
                     x=c(        0,    0.2,    0.4,   0.4,   0.35,   0.55,   0.38, 1.0),
                     y=c(     -0.03,    0.2,   -0.08,   0.2,  0.03,   0.07,  -0.2, 0))
dagified <- dagify( INTR ~ REM + COMP + ADD + DUP + TEAM, 
                    COMP ~ TEAM + DUP,
                    REM ~ TEAM + CLEAN,
                    ADD ~ COMP + CLEAN + KNOW + DUP,
                    DUP ~ TEAM,
                    CLEAN ~ TEAM,
                    KNOW ~ TEAM + DUP + COMP,
                    exposure = "TEAM",
                    outcome = "INTR",
                    latent = c("CLEAN", "KNOW"),
                    coords=coords)
(p <- ggdag_status(dagified) + theme_dag() + theme(legend.position="bottom") + guides(color = guide_legend(override.aes = list(size = 2), title=NULL, direction="horizontal", nrow = 1)) 
  )

```
```{r}
figsave("causal-dag.pdf", p)
```

