---
title: "Visualizing the Bayesian Workflow With BRMS"
author: "Anders Sundelin"
date: "2022-12-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rstan)
library(brms)
library(bayesplot)
library(loo)
library(tidybayes)
```

## Visualizing the Bayesian Workflow

Monica Alexander has a great tutorial on prior predictive checks at https://www.monicaalexander.com/posts/2020-28-02-bayes_viz/.

```{r generate-sample-data, eval=FALSE}

# download better done through your browser, it will have a longer timeout than Rs default 60 seconds...
#download.file("https://data.nber.org/natality/2017/natl2017.csv.zip", destfile = "natl2017.csv.zip")

# read in and select
d <- read_csv("~/Downloads/natl2017.csv.zip")
head(d)
d <- d %>%
  select(mager, mracehisp, meduc, bmi, sex, combgest, dbwt, ilive)

# only use a sample of the births
set.seed(853)
ds <- d[sample(1:nrow(d), nrow(d)*0.001),]

ds <- ds %>% mutate(dbwt= dbwt/1000)
write_rds(ds, path = "births_2017_sample.RDS")
```

```{r ingest}
ds <- read_rds("births_2017_sample.RDS")
head(ds)
```
## Simple Exploratory Data Analysis

```{r}
ds <- ds %>% 
  rename(birthweight = dbwt, gest = combgest) %>% 
  mutate(preterm = ifelse(gest<32, "Y", "N")) %>% 
  filter(ilive=="Y",gest< 99, birthweight<9.999)
ds %>% 
  ggplot(aes(log(gest), log(birthweight))) + 
  geom_point() + geom_smooth(method = "lm") + 
  scale_color_brewer(palette = "Set1") + 
  theme_bw(base_size = 14) +
  ggtitle("birthweight v gestational age")
```

Separating out premature babies.

```{r}
ds %>% 
  ggplot(aes(log(gest), log(birthweight), color = preterm)) + 
  geom_point() + geom_smooth(method = "lm") + 
  scale_color_brewer(palette = "Set1") + 
  theme_bw(base_size = 14) + 
  ggtitle("birthweight v gestational age")
```


## Candidate Models

Alternative 1 - simple model, straight relationship between log birth weight and log gestational age:

$log(y_i) \sim N(\beta_0 + \beta_1 log(x_i), \sigma^2) $

Alternative 2 has an interaction term (both fixed effect and effect proportional to the original log gestational age)

$log(y_i) \sim N(\beta_0 + \beta_1 log(x_i) + \gamma_0 z_i + \gamma_1 log(x_i) z_i, \sigma^2) $

where 

* $y_i$ is weight in kg
* $x_i$ is gestational age in weeks
* $z_i$ is preterm (0 or 1, if gestational age is less than 32 weeks)



## Prior Predictive Checks

We do simulations from priors and likelihood and plot the resulting distribution, to see whether we come up with reasonable values.

### Vague priors


```{r vague}
set.seed(182)
nsims <- 100
sigma <- 1 / sqrt(rgamma(nsims, 1, rate = 100))
beta0 <- rnorm(nsims, 0, 100)
beta1 <- rnorm(nsims, 0, 100)

# scale and center outcome (comes from the real data)
dsims <- tibble(log_gest_c = (log(ds$gest)-mean(log(ds$gest)))/sd(log(ds$gest)))

# build the parameter, based on the simulated priors (sigma, beta0/beta1 above)
for(i in 1:nsims){
  # calculate new mean value
  this_mu <- beta0[i] + beta1[i]*dsims$log_gest_c 
  # create a new column, with as many rows as dsims, and contents of new mean value, plus or minus a value corresponding to the stddev
  dsims[paste0(i)] <- this_mu + rnorm(nrow(dsims), 0, sigma[i])
}

dsl <- dsims %>% 
  pivot_longer(`1`:`100`, names_to = "sim", values_to = "sim_weight")

dsl %>% 
  ggplot(aes(sim_weight)) + geom_histogram(aes(y = ..density..), bins = 20, fill = "turquoise", color = "black") + 
  xlim(c(-1000, 1000)) + 
  geom_vline(xintercept = log(104), color = "purple", lwd = 1.2, lty = 2) + 
  theme_bw(base_size = 16) + 
  annotate("text", x=300, y=0.0022, label= "Anders'\ncurrent weight", 
           color = "purple", size = 5) 
```

Clearly, babies are not that heavy --- they should be in the 1.0-5.0 range (perhaps stretching up to 8-9 kg at the most, or in some cases perhaps lower than 1.0 kg)

### Weakly Informative Priors

```{r}
sigma <- abs(rnorm(nsims, 0, 1))
beta0 <- rnorm(nsims, 0, 1)
beta1 <- rnorm(nsims, 0, 1)

dsims <- tibble(log_gest_c = (log(ds$gest)-mean(log(ds$gest)))/sd(log(ds$gest)))

for(i in 1:nsims){
  this_mu <- beta0[i] + beta1[i]*dsims$log_gest_c 
  dsims[paste0(i)] <- this_mu + rnorm(nrow(dsims), 0, sigma[i])
}

dsl <- dsims %>% 
  pivot_longer(`1`:`100`, names_to = "sim", values_to = "sim_weight")

dsl %>% 
  ggplot(aes(sim_weight)) + geom_histogram(aes(y = ..density..), bins = 20, fill = "turquoise", color = "black") + 
  geom_vline(xintercept = log(104), color = "purple", lwd = 1.2, lty = 2) + 
  theme_bw(base_size = 16) + 
  annotate("text", x=8, y=0.2, label= "Anders'\ncurrent weight", color = "purple", size = 5)
```

### Running the models from Stan

Skipped

Formatting the data for the models

```{r}
ds$log_weight <- log(ds$birthweight)
ds$log_gest_c <- (log(ds$gest) - mean(log(ds$gest)))/sd(log(ds$gest))

N <- nrow(ds)
log_weight <- ds$log_weight
log_gest_c <- ds$log_gest_c 
preterm <- ifelse(ds$preterm=="Y", 1, 0)

# put into a list
stan_data <- list(N = N,
                  log_weight = log_weight,
                  log_gest = log_gest_c, 
                  preterm = preterm)
```

```{r}
mod1 <- stan(data = stan_data, 
             file = "models/simple_weight.stan",
             iter = 500,
             seed = 243)

mod2 <- stan(data = stan_data, 
             file = "models/simple_weight_preterm_int.stan",
             iter = 500,
             seed = 263)
```
```{r}
summary(mod1)[["summary"]][c(paste0("beta[",1:2, "]"), "sigma"),]
```

```{r}
summary(mod2)[["summary"]][c(paste0("beta[",1:4, "]"), "sigma"),]
```

### Running the Models In BRMS

Why are we not using our priors? Instead, we are using BRMS's deafult priors. 

```{r}
mod1b <- brm(log_weight~log_gest_c, data = ds)
mod2b <- brm(log_weight~log_gest_c*preterm, data = ds)
```

## Posterior Predictive Checks

```{r}
set.seed(1856)
y <- log_weight
yrep1 <- rstan::extract(mod1)[["log_weight_rep"]]
samp100 <- sample(nrow(yrep1), 100)
ppc_dens_overlay(y, yrep1[samp100, ])  
```

```{r}
yrep2 <- rstan::extract(mod2)[["log_weight_rep"]]
ppc_dens_overlay(y, yrep2[samp100, ])  
```

Manual handling of the density plots:

```{r}
# first, get into a tibble
rownames(yrep1) <- 1:nrow(yrep1)
dr <- as_tibble(t(yrep1))
dr <- dr %>% bind_cols(i = 1:N, log_weight_obs = log(ds$birthweight))

# turn into long format; easier to plot
dr <- dr %>% 
  pivot_longer(-(i:log_weight_obs), names_to = "sim", values_to ="y_rep")

# filter to just include 100 draws and plot!
dr %>% 
  filter(sim %in% samp100) %>% 
  ggplot(aes(y_rep, group = sim)) + 
  geom_density(alpha = 0.2, aes(color = "y_rep")) + 
  geom_density(data = ds %>% mutate(sim = 1), 
               aes(x = log(birthweight), col = "y")) + 
  scale_color_manual(name = "", 
                     values = c("y" = "darkblue", 
                                "y_rep" = "lightblue")) + 
  ggtitle("Distribution of observed and replicated birthweights") + 
  theme_bw(base_size = 16)
```

```{r}
ppc_stat(log_weight, yrep1, stat = 'median')
```

```{r}
ppc_stat(log_weight, yrep2, stat = 'median')
```

Replicating the proportion of births less than 2.5 kg.

```{r}
t_y <- mean(y<=log(2.5))
t_y_rep <- sapply(1:nrow(yrep1), function(i) mean(yrep1[i,]<=log(2.5)))
t_y_rep_2 <- sapply(1:nrow(yrep2), function(i) mean(yrep2[i,]<=log(2.5)))

ggplot(data = as_tibble(t_y_rep), aes(value)) + 
    geom_histogram(aes(fill = "replicated")) + 
    geom_vline(aes(xintercept = t_y, color = "observed"), lwd = 1.5) + 
  ggtitle("Model 1: proportion of births less than 2.5kg") + 
  theme_bw(base_size = 16) + 
  scale_color_manual(name = "", 
                     values = c("observed" = "darkblue"))+
  scale_fill_manual(name = "", 
                     values = c("replicated" = "lightblue")) 
```

```{r}
ggplot(data = as_tibble(t_y_rep_2), aes(value)) + 
    geom_histogram(aes(fill = "replicated")) + 
    geom_vline(aes(xintercept = t_y, color = "observed"), lwd = 1.5) + 
  ggtitle("Model 2: proportion of births less than 2.5kg") + 
  theme_bw(base_size = 16) + 
  scale_color_manual(name = "", 
                     values = c("observed" = "darkblue"))+
  scale_fill_manual(name = "", 
                     values = c("replicated" = "lightblue")) 
```

## Posterior Predictions with brms

```{r}
yrep1b <- posterior_predict(mod1b)
ds_yrep1 <- ds %>% 
  select(log_weight, log_gest_c) %>% 
  add_predicted_draws(mod1b)
head(ds_yrep1)
```

```{r}
pp_check(mod1b, type = "dens_overlay", nsamples = 100)
```

```{r}
pp_check(mod1b, type = "stat", stat = 'median', nsamples = 100)
```

```{r}
loglik1 <- rstan::extract(mod1)[["log_lik"]]
loglik2 <- rstan::extract(mod2)[["log_lik"]]
loo1 <- loo(loglik1, save_psis = TRUE)
loo2 <- loo(loglik2, save_psis = TRUE)
```

```{r}
loo_compare(loo1, loo2)
```

```{r}
plot(loo1)
```

Plotting the LOO Probability Integral Transform (LOO-PIT).
Looks to see where each point $y_i$ falls in its predictive distribution $p(y_i|y_{-i})$
If the model is well calibrated, these should look like uniform distributions (spread evenly between 0 and 1).

```{r}
ppc_loo_pit_overlay(yrep = yrep1, y = y, lw = weights(loo1$psis_object)) + ggtitle("LOO-PIT Model 1")
```

```{r}
ppc_loo_pit_overlay(yrep = yrep2, y = y, lw = weights(loo2$psis_object)) + ggtitle("LOO-PIT Model 2")
```

## LOO-CV with brms output

```{r}
loo1b <- loo(mod1b, save_psis = TRUE)
loo2b <- loo(mod2b, save_psis = TRUE)
loo_compare(loo1b, loo2b)
```

```{r}
plot(loo1b)
```

```{r}
ppc_loo_pit_overlay(yrep = yrep1b, y = y, lw = weights(loo1b$psis_object)) + ggtitle("LOO-PIT BRMS Model 1")

```

```{r}
yrep2b <- posterior_predict(mod2b)

ppc_loo_pit_overlay(yrep = yrep2b, y = y, lw = weights(loo2b$psis_object)) + ggtitle("LOO-PIT BRMS Model 2")

```

